Portfolio conservation of exploited metapopulations\\Portfolio prioritization of salmon conservation

### 2013-06-05 inst/ms/ms.tex
Make a variety of ms edits

The management of financial portfolios provides another way of
considering risk. Economists consider the risk and performance of a
financial portfolio as a function of the weighting of individual assets
that make up the portfolio. Modern Portfolio Theory (MPT) proposes that
there is a set of portfolios that maximizes expected return for a level
of expected risk or minimizes expected risk for a level of expected
return \citep{Markowitz1952, Markowitz1959}. This optimal set contains
portfolios that range along a continuum of risk-tolerance; economists
refer to this set as the efficient frontier.

Portfolios provide another way of managing risk - explain mean-variance
optimization and how this translates to ecology
\citep{Figge2004}\citep{Hoekstra2012}
\citep[\citet{Ando2012}]{Ando2011}\citep[\citet{Markowitz1959}]{Markowitz1952}

We use salmon as an example, describe the components of the example,
provide references on metapopulation existence \citep{Schindler2010}

@@

\subsection{Overview}

We developed a salmon metapopulation simulation model that includes a
stock-recruit relationship, demographic stochasticity, straying between
populations, varying responses to the environment, escapement target
setting, and implementation uncertainty. Under two kinds of
environmental regimes we tested different conservation rules of thumb
and evaluated these plans in risk-return space similar to how financial
managers evaluate financial portfolios. See Figure \ref{f:sim-flow} for
an illustration of the overall simulation structure.

@@

In our ecological portfolio, we defined assets as stream-level
populations and the portfolio as a salmon metapopulation (Table
\ref{t:port}). The portfolio investors were the stakeholders in the
fishery and metapopulation performance. For example, we could consider
fisheries managers, conservation agencies, or first Nations groups as
investors. We defined asset value as abundance (returns) of stream
populations and value of the portfolio as the overall metapopulation
abundance. In this scenario, the equivalent to financial rate of return
is the generation-to-generation rate of change of metapopulation
abundance. We defined the financial asset investment weights as the
carrying capacity of the stream populations, specifically the unfinished
equilibrium stock size.

@@

The salmon metapopulation dynamics in our simulation were governed by a
spawner-return relationship with stochastic noise and straying between
populations.

@@

We defined the spawner-return relationship with a Ricker model
\citep{Ricker1954},

\begin{equation}
R_{ti} = S_{ti}e^{a_{ti}(1-S_{ti}/b_i) + w_{ti}},
\end{equation}

@@

@@

We implemented straying as in \citet{Cooper1999}. We set up the
metapopulation in a simple scenario: the populations were arranged in a
line and those that were nearer to each other were more likely to stray
between each other. Two parameters controlled the straying: the fraction
of fish $f_{\mathrm{stray}}$ that stray from their natal stream in any
generation and the rate $m$ at which this straying between streams
decays with distance. We calculated the number of salmon straying from
stream $j$ to stream $i$ as,

\begin{equation}
  \mathrm{strays}_{tij} = f_{\mathrm{stray}} R_{tj}
    \frac{e^{-m \lvert i-j \rvert }}
      {\displaystyle\sum\limits_{\substack{k = 1 \\
    k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
  \label{eq:stray}
\end{equation}

@@

Our simulation used a simple set of rules to establish escapement
targets and implement harvesting. Every 5(TODO CAN VARY) years our
simulation fitted a spawner-return function and target harvest rate
$H_{\mathrm{tar}}$ was set based on \citet{Hilborn1992} as

@@

\noindent
where $\alpha_h$ and $\beta_h$ are the location and shape parameters in
a beta distribution. They can be calculated from the desired mean
$\mu_h$ and standard deviation $\sigma_h$ as: TODO ADD REFERENCES

\begin{align}
  \alpha_h &= \frac{\mu_h^2(1 - \mu_h)}{\sigma_h - 1/\mu_h}\\
   \beta_h &= \frac{\alpha}{\mu_h - 1}.
\end{align}

@@

We can generally break environmental dynamics down into short-term and
long-term changes. We evaluated the contribution of the environment to
metapopulation performance under these two components separately.

short-term forcing: stationary AR1; long-term forcing: linear increase

@@

We evaluated two sets of conservation rules of thumb: (1) spatial
response diversity rules of thumb in an idealistic situation where you
can detect response diversity, and (2) a more realistic scenario where
we know little about response diversity and we're left with a choice of
how many populations to conserve. We then considered how optimal these
strategies are across all possible conservation-prioritization
strategies.

@@

We evaluated four spatial conservation rules of thumb (Figure
\ref{f:curves}). In all scenarios, four populations are conserved and
the unfished equilibrium biomass of the remaining populations is reduced
to near elimination (five salmon). These reduced populations can
therefore still receive straying salmon but are unlikely to re-build on
their own to any substantial volume of salmon.

The four scenarios are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Conserve an even sampling of response diversity. In other words,
  conserve alternating subpopulations.
\item
  Conserve the most stable populations only.
\item
  Conserve one half of the metapopulation.
\item
  Conserve the other half of the metapopulation.
\end{enumerate}

@@

In reality we rarely know precise levels of response diversity. We
therefore additionally considered a case where the conservation is
randomly assigned with respect to response diversity but where different
numbers of streams can be conserved. We considered conserving from two
to 16 streams. Similarly to the spatial strategies, we reduced the
carrying capacity of the remaining streams to the nominal level of five
salmon.

@@

To test how optimal the strategies are, we generated metapopulation
portfolios in which the investment weights (Ricker $b$ parameters) were
randomly assigned. This Monte-Carlo strategy illustrates a large sample
of possible portfolios in risk-return space. We then placed the
conservation strategies listed above into this space to determine how
close to the efficient frontier they lie.

@@

Fig. \ref{f:eg-n-linear}

Fig. \ref{f:eg-n-arma}

Fig. \ref{f:eg-sp-arma}

Fig. \ref{f:eg-sp-linear}

\subsection{Base case}

\subsection{Effect of X on \ldots{}}

\subsection{Effect of Y on \ldots{}}

@@

\section{References}

@@

\begin{table}[h!]
\centering
\small
\caption{Salmon metapopulation parameters with base case and alternate values.}
\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
\toprule
Description                                                      & Parameter             & Base [lower, upper] \\
\midrule
Stock-recruit residual standard deviation (on log scale)         & $\sigma_v$            & 0.30 [0.05, 0.50] \\
First order (AR1) serial correlation of stock-recruit residuals  & $\rho_w$              & 0.40 [0, 0.80] \\
Fraction of fish that stray from natal streams                   & $f_{\mathrm{stray}}$  & 0.02 [0, 0.10] \\
Exponential rate of decay of straying with distance              & $m$                   & 0.3 [0.05, 0.5] \\
Standard deviation of beta distribution for implementation error & $\sigma_{h}$          & 0.05 [0, 0.20] \\
Frequency of assessment (years)                                  & $f_{\mathrm{assess}}$ & 20 [5, 50] \\
\bottomrule
\end{tabular}
\label{tab:salm-pars}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics[height=5.5in]{../examples/simulation-diagram2.pdf}
\caption{Flow chart of the salmon-metapopulation simulation. There are $n$ 
salmon populations and $t$ generations. Blue text indicates values that are 
generated before the simulation progresses through time. Red text indicates 
steps in which calculations are performed. Black text indicates values that are 
calculated. Grey text indicates parameters that can be set. Green text 
indicates the looping structure of the simulation.}
\label{f:sim-flow}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
\caption{Different ways of prioritizing response-diversity conservation. The 
coloured lines represent environmental-tolerance curves for conserved 
populations. The colours identify populations that thrive at low (cool colours) 
to high (warm colours).}
\label{f:curves}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=5.0in]{../examples/quasi-extinct.pdf}
\caption{First draft of a quasi-extinction figure. The y-axis shows the 
cumulative proportion of subpopulations that have had abundance dip below a 
threshold of 20\% of unfished equilibrium abundance. We could show a (binomial) 
measure of uncertainty around these if we want to. Some interesting points: the 
cumulative proportion of ``quasi-extinction'' is higher for both ``one half'' 
scenarios. If it's set up correctly, then this means that there's a benefit to 
the subpopulations to investing in a more ``optimal'' portfolio. The }
\label{f:n-mv}
\end{figure}

>>>>

The management of financial portfolios provides another way of
considering risk. Economists consider the risk and performance of a
financial portfolio as a function of the weighting of individual assets
that make up the portfolio. Modern Portfolio Theory (MPT) proposes that
there is a set of portfolios that maximizes expected return for a level
of expected risk or minimizes expected risk for a level of expected
return \citep{markowitz1952, markowitz1959}. This optimal set contains
portfolios that range along a continuum of risk-tolerance; economists
refer to this set as the efficient frontier.

Portfolios provide another way of managing risk - explain mean-variance
optimization and how this translates to ecology
\citep{figge2004}\citep{hoekstra2012}
\citep[\citet{ando2012}]{ando2011}\citep[\citet{markowitz1959}]{markowitz1952}

We use salmon as an example, describe the components of the example,
provide references on metapopulation existence \citep{schindler2010}

@@

We developed a salmon metapopulation simulation model that includes a
stock-recruit relationship, demographic stochasticity, straying between
populations, varying responses to the environment, escapement target
setting, and implementation uncertainty. Under two kinds of
environmental regimes we tested different conservation rules of thumb
and evaluated these plans in risk-return space similar to how financial
managers evaluate financial portfolios. See Figure \ref{f:sim-flow} for
an illustration of the overall simulation structure. We provide a
package \texttt{metafolio} for the statistical software \texttt{R}
\citep{r2013} as an appendix, which can be used to carry out the
simulations and analyses we describe in this paper.

@@

In our ecological portfolios, we defined assets as stream-level
populations and the portfolios as salmon metapopulations (Table
\ref{t:port}). We use the terms \textit{stream} and \textit{populations}
interchangeably to represent the portfolio assets. The portfolio
investors were the stakeholders in the fishery and metapopulation
performance. For example, we could consider fisheries managers,
conservation agencies, or First Nations groups as investors. We defined
asset value as the abundance of returning salmon in each stream and
value of the portfolio as the overall metapopulation abundance. In this
scenario, the equivalent to financial rate of return is the
generation-to-generation rate of change of metapopulation abundance. We
defined the financial asset investment weights as the carrying capacity
of the stream populations --- specifically the unfinished equilibrium
stock size --- since maintaining or restoring habitat requires money,
time, and resources.

@@

The salmon metapopulation dynamics in our simulation were governed by a
spawner-return relationship with stochastic noise and by straying
between populations.

@@

We defined the spawner-return relationship with a Ricker model
\citep{ricker1954},

\begin{equation}
R_{ti} = S_{ti}e^{a_{ti}(1-S_{ti}/b_i) + w_{ti}}
\end{equation}

@@

We manipulated the capacity and productivity parameters $b$ and $a$ as
part of the portfolio simulation. The capacity parameters $b_i$ were
controlled by the investment weights in the populations. For example, a
large investment in a stream was represented by a larger unfished
equilibrium stock size $b$ for stream $i$. The productivity parameters
$a_{ti}$ were controlled by the interaction between an environmental
signal and the stream-level population environmental tolerance curves.
The environmental tolerance curves were generated as parabolas according
to

\begin{equation}
  a_{ti} =
  \begin{cases}
    W_i (e_t - e_i^{\mathrm{opt}})^2 + a_i^{\mathrm{max}},
      & \text{if } a_{ti} > 0\\
      0, & \text{if } a_{ti} \leq 0
  \end{cases}
\end{equation}

\noindent where $W_i$ controls the width of the curve for population
$i$, $e_t$ represents the environmental value at generation $t$,
$e_i^{\mathrm{opt}}$ represents the optimal environmental value for
population $i$, and $a_i^{\mathrm{max}}$ represents the maximum possible
$a$ value for population $i$. See Figure \ref{f:all-curves} for example
environmental tolerance curves.

@@

We implemented straying as in \citet{cooper1999}. We set up the
metapopulation in a simple scenario: we arranged the populations in a
line and those that were nearer to each other were more likely to stray
between each other. Two parameters controlled the straying: the fraction
of fish $f_{\mathrm{stray}}$ that stray from their natal stream in any
generation and the rate $m$ at which this straying between streams
decays with distance. We calculated the number of salmon straying from
stream $j$ to stream $i$ as

\begin{equation}
  \mathrm{strays}_{tij} = f_{\mathrm{stray}} R_{tj}
    \frac{e^{-m \lvert i-j \rvert }}
      {\displaystyle\sum\limits_{\substack{k = 1 \\
    k \neq j}}^{n} e^{-m \lvert k-j \rvert }}
  \label{eq:stray}
\end{equation}

@@

Our simulation used a simple set of rules to establish escapement
targets and harvest the fish. Every $f_\mathrm{assess}$ years (base case
of five years) our simulation fitted a spawner-return function and
target harvest rate $H_{\mathrm{tar}}$ was set based on
\citet{hilborn1992} as

@@

\noindent where $\alpha_h$ and $\beta_h$ are the location and shape
parameters in a beta distribution. They can be calculated from the
desired mean $\mu_h$ and standard deviation $\sigma_h$ as
\citep[p.~97]{morgan1990}

\begin{align}
  \alpha_h &= \mu_h^2
                \left(
                \frac{1 - \mu_h}{\sigma_h^2} - \frac{1}{\mu_h}
                \right)\\
   \beta_h &= \alpha \left({\frac{1}{\mu_h} - 1}\right).
\end{align}

@@

We evaluated portfolio performance under short- and long-term
environmental dynamics. We represented short-term dynamics as a
stationary first-order autoregressive process, AR(1), with correlation
$\rho_e$

\begin{equation}
  e_t = e_{t-1} \rho_e + d_t, d_t \sim \mathrm{N}(0, \sigma_d)
\end{equation}

\noindent where $e_t$ represents the environmental value in generation
$t$ and $d$ represents normally distributed deviations of mean 0 and
standard deviation $\sigma_d$. We represented long-term environmental
dynamics as a linear shift in the environmental value through time

\begin{equation}
  e_t = \alpha_e + \beta_e t
\end{equation}

\noindent where $\alpha_e$ represents the intercept and $\beta_e$ the
slope. To maintain a balanced response, we centered the trend so that
midway through the simulation (after any burn-in period) the
environmental value was at the mean environmental tolerance.

@@

We evaluated two sets of conservation rules of thumb: (1) spatial
response diversity rules of thumb in an idealistic scenario where you
can detect response diversity, and (2) a more realistic scenario where
we know little about response diversity and we're left with a choice of
how many populations to conserve. We then considered how optimal these
strategies were across all possible conservation-prioritization
strategies.

@@

We evaluated four spatial conservation rules of thumb (Figure
\ref{f:curves}). In all spatial scenarios, we conserved four populations
and set the unfished equilibrium biomass of the remaining populations to
near elimination (five salmon). These reduced populations could
therefore still receive straying salmon but were unlikely to re-build on
their own to a substantial abundance. The four scenarios were:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Conserve an even sampling of response diversity. In other words,
  conserve alternating subpopulations.
\item
  Conserve the most stable populations only.
\item
  Conserve one half of the metapopulation.
\item
  Conserve the other half of the metapopulation.
\end{enumerate}

@@

In reality we rarely know precise levels of response diversity. We
therefore additionally considered a case where the conservation was
randomly assigned with respect to response diversity but where different
numbers of streams can be conserved. We considered conserving from two
to 16 streams. Similarly to the spatial strategies, we reduced the
carrying capacity of the remaining streams to the nominal level of five
salmon.

@@

To test how optimal the strategies were, we generated metapopulation
portfolios in which the investment weights (Ricker $b$ parameters) were
randomly assigned such that the total available ``investment'' remained
constant. This Monte-Carlo strategy illustrates a large sample of
possible portfolios in risk-return space. We then placed the
conservation strategies listed above into this space to determine how
close to the efficient frontier they lie. {[}Figure not shown yet{]}

@@

Figure \ref{f:curves} Figure \ref{f:sp-mv} Figure \ref{f:n-mv}

Figure \ref{f:eg-n-linear} Figure \ref{f:eg-n-arma} Figure
\ref{f:eg-sp-arma} Figure \ref{f:eg-sp-linear}

@@

@@

\begin{table}[h!]
\centering
\small
\caption{Salmon metapopulation parameters with base case and alternate values.}
\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
\toprule
Description                                                      & Parameter             & Base [lower, upper] \\
\midrule
Stock-recruit residual standard deviation (on log scale)         & $\sigma_v$            & 0.30 [0.05, 0.50] \\
First order AR(1) serial correlation of stock-recruit residuals  & $\rho_w$              & 0.40 [0, 0.80] \\
Fraction of fish that stray from natal streams                   & $f_{\mathrm{stray}}$  & 0.02 [0, 0.10] \\
Exponential rate of decay of straying with distance              & $m$                   & 0.3 [0.05, 0.5] \\
Standard deviation of beta distribution for implementation error & $\sigma_{h}$          & 0.05 [0, 0.20] \\
Frequency of assessment (years)                                  & $f_{\mathrm{assess}}$ & 5 [1, 20] \\
Environmental tolerance parameters ... & & \\
\bottomrule
\end{tabular}
\label{t:pars}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics[height=5.5in]{../examples/simulation-diagram2.pdf}
\caption{Flow chart of the salmon-metapopulation simulation. There are $n$ salmon populations and $t$ generations. Blue text indicates values that are generated before the simulation progresses through time. Red text indicates steps in which calculations are performed through time. Black text indicates values that are calculated. Grey text indicates parameters that can be set. Green text indicates the looping structure of the simulation.}
\label{f:sim-flow}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=3.0in]{../examples/thermal-curve-scenarios.pdf}
\caption{Different ways of prioritizing response-diversity conservation. Panel (a) shows the thermal tolerance cures for ten possible populations and panels (b--e) show different ways of prioritizing four of those populations. The curves describe how productivity varies with the environment for a given population. Some populations thrive at low environmental values (cool colours) and some at high (warm colours) values. Some are tolerant to a wider range of environmental conditions (yellow-to-green colours) but with a lower maximum productivity. The total possible productivity across environmental values (the area under the curves) is the same.}
\label{f:curves}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=5.0in]{../examples/quasi-extinct.pdf}
\caption{First draft of a quasi-extinction figure I THINK THIS WILL BE 
SCRAPPED OR CHANGED. The y-axis shows the cumulative proportion of 
subpopulations that have had abundance dip below a threshold of 20\% 
of unfished equilibrium abundance. We could show a (binomial) 
measure of uncertainty around these if we want to. Some interesting points: the 
cumulative proportion of ``quasi-extinction'' is higher for both ``one half'' 
scenarios. If it's set up correctly, then this means that there's a benefit to 
the subpopulations to investing in a more ``optimal'' portfolio. The }
\label{f:ext}
\end{figure}
